Genome reconstruction in R
========================================================

The purpose of this small guide is to create a reproducible workflow for generating layouts of genomes.

### Load needed packages
```{r load_packages}
library(Biostrings)
```

### Load original scaffolds
The original scaffolds can be loaded directly as a fasta file.
```{r load_org_scaffolds}
org <- readDNAStringSet("scaffolds.fa", format = "fasta")
```

Access a specific sequence.
```{r access_sequences}
seq1 <- org[1]
seq1
```

Reverse complement.
```{r reverse_complement}
seq1_rc <- reverseComplement(seq1)
seq1_rc
```

Join two sequences.
```{r join}
seqNew <- xscat(seq1, seq1_rc)
seqNew
```

Extract a subset of the nucleotides from a given sequence.
```{r split}
seqNew_sub <- subseq(seqNew, start=1, end=15)
seqNew_sub
```

Join several sequences in a new object.
```{r create object}
new_list <- c(seqNew, seq1_rc, seq1)
names(new_list) <- c("seqA", "seqB", "seqC")
new_list
```

Write the sequences to a file.
```{r write}
writeXStringSet(new_list, "out.fa")
```

