Download MG-RAST metagenome data
========================================================

## Load packages

```{r libraries}
library("reshape2")
```

## Specify a metagenome
```{r metagenome}
mg <- "4463936.3"
```

# Download data

## Download predicted proteins

```{r dl proteins}
paste("http://api.metagenomics.anl.gov/1/download/mgm",mg,"?file=350.1", sep="")
```

## Download cluster information

```{r cluster}
paste("http://api.metagenomics.anl.gov/1/download/mgm",mg,"?file=550.4", sep="")
```

## Download protein annotation information

```{r annotation}
paste("http://api.metagenomics.anl.gov/1/download/mgm",mg,"?file=650.3", sep="")
```

## Download ontology information

```{r ontology}
paste("http://api.metagenomics.anl.gov/1/download/mgm",mg,"?file=650.2", sep="")
```

# Read data

## Protein identifiers

```{r protein id}
p <- readLines(con="350.genecalling.coding.faa")
pid <- as.data.frame(grep(">", p, value = T))
colnames(pid) <- "pid"
pid$pid <- as.character(pid$pid)
pid[1:2,]

pid <- as.data.frame(sub(">", "", pid$pid))
colnames(pid) <- "pid"
rm(p)
```

## Cluster identifiers

```{r cluster id}
c <- read.delim("550.cluster.aa90.mapping", header = F)
colnames(c) <- c("cid", "seed", "add", "procent")

c2 <- cbind.data.frame(c$cid,paste(c$seed, c$add, sep = ","))
colnames(c2) <- c("cid","pid")

myformat <- function(x){
  t2 <-strsplit(as.character(x[2]), ",")
  out <- cbind(x[1],as.data.frame(t2))
  colnames(out) <- c("cid","pid")
  return(out)
}

cid.list <- apply(c2, 1, myformat)
cid <- melt(cid.list, id.vars=c("cid","pid"))
cid <- cid[,1:2]
rm(cid.list)
rm(c)
rm(c2)
```

## Merge cluster id with protein id

```{r merge pid and cid}
pc <- merge(x = pid, y = cid, by = "pid", all.x = T)

pc$pid <- as.character(pc$pid)
pc$cid <- as.character(pc$cid)

myreplace <- function(x){
  if (is.na(x[2])){
    x[2] <- x[1]
  }
  return(x)
}

pc <- as.data.frame(t(apply(pc, 1, myreplace)))
colnames(pc) <- c("pid", "id")
```

## Read ontology data 

```{r read ontology data}
o <- read.delim("650.superblat.expand.ontology", header = F)
colnames(o) <- c("md5","ps", "sim","aln","eval", "annotation", "oid", "db")

o <- subset(o, db == "Subsystem" | db == "COG")
```

## Read protein data

```{r read protein data}
p <- read.delim("650.superblat.expand.protein", header = F)
colnames(p) <- c("md5","ps", "sim","aln","eval", "annotation", "species", "db", "extra")

p <- subset(p, db == "RefSeq" | db == "SEED")
```

## Export annotations

```{r export annotations}
e <- subset(p, db == "RefSeq" & eval < 1e-5)
edp <- e[!duplicated(e$ps),]

export <- merge(x = pc, y = edp[,c("ps", "annotation","species")], by.x = "id", by.y = "ps", all.x = T)

export.header <- paste(">",export$pid," ",export$annotation, " [", export$species,"]", sep = "")

write.table(export.header, file = "header.txt", row.names = F , col.names=F, quote=F)
```

## Typical workflow

You have a list of identifiers.
```{r workflow}



```

