Network analysis
========================================================

How to analyse PE connection networks in R.

```{r load packages}
library("igraph")
library("RColorBrewer")
library("ggplot2")
```

```{r define color palette}
gr<-colorRampPalette(c("green", "red"))
palette(gr(100))
```


Read all data.

```{r load data}
PE<-read.delim("connections.txt", header =T)
meta<-read.delim("metadata.txt", header= T)
```

Generate graph.

```{r generate graph}
PEg<-graph.data.frame(PE[,c(1,3,4)], directed = F)
PEg
```

Sort the metadata to match the order of the graph.

```{r sort meta}
meta<-meta[match(V(PEg)$name,meta$Scaffold),]
```

Add attributes to `Vertices`.

```{r add to vertrices}
V(PEg)$label<- V(PEg)$name
V(PEg)$label.cex <- .7
V(PEg)$label.color <- "black"
V(PEg)$label.font <- 2
V(PEg)$frame.color<-"black"
V(PEg)$size <- sqrt(meta$Length)/25 + 2
V(PEg)$color <- (log(meta$Coverage)-min(log(meta$Coverage)))*(98/(max(log(meta$Coverage)-min(log(meta$Coverage)))))+1
V(PEg)$length <- meta$Length
V(PEg)$coverage <- meta$Coverage
V(PEg)$degree <- degree(PEg)
```

Add attributes to `Edges`.

```{r add to edges}
E(PEg)$color <- "black"
E(PEg)$width <- log(E(PEg)$connections)/2
```

## Usefull things to do with the graph

### Clusters
Overview of clusters in the network graph and their size.

```{r clusters}
PEgc<-clusters(PEg)
PEgc
```

### Standard plot
Plot the graph.

```{r standard plot, fig.width=10, fig.height=10}
layout1 <- layout.fruchterman.reingold(PEg)
plot(PEg, layout=layout1)
```

### Find scaffolds with too many connections

First add the number of connections each scaffold have to the metadata.
```{r explore connections}
meta<-cbind(meta, V(PEg)$degree)
colnames(meta)[4] <- "Degree"
```

Spotting repeats by looking at connections versus repeats.
```{r plot degree}
ggplot(meta, aes(x = Coverage, y = Degree)) +
  geom_point(size = 2)
```

Looking in detail at scaffolds with repeats integrated in ends. Normal scaffolds should have max 2 connections.
```{r plot degree bad}
ggplot(meta, aes(x = Coverage, y = Degree)) +
  geom_text(label = meta$Scaffold) +
  xlim(0, 400) +
  ylim(0,10)
```

Printing the scaffolds that needs to be looked at.
```{r report bad scaffolds}
subset(meta, Coverage < 400 & Degree > 2)
```

### Subset the graph

Using attributes to subset the graph.
```{r subset}
bad.vs<-V(PEg)[V(PEg)$coverage > 400] 
PEgs<-delete.vertices(PEg, bad.vs)
```

Plot the subset.

```{r plot subset, fig.width=10, fig.height=10}
layout1 <- layout.fruchterman.reingold(PEgs)
plot(PEgs, layout=layout1)
```
