Create coverage plots from BAM files
========================================================

Load the needed packages. The Rsamtools package can be used to interface with BAM files, which are used to store read mappings.
```{r load packages, message=F}
#source("http://bioconductor.org/biocLite.R")
#biocLite("Rsamtools")
library("Rsamtools")
library("ggplot2")
```

Load the gene positions within each sequenced instert. 
```{r load_gene}
gpos <- read.delim("data/gene.position.txt", header = T)
```

Making a function that extracts the covrage profile for a specific reference sequence in a BAM file. The BAM files contains all mappings (i.e. to all different inserts) for each sample, but we need to extract the mappings to a particular sequenced insert. The BAM file needs to be sorted and indexed.
```{r function}
extcov <- function(bam, index, gene, start=1, end=10000){
  which <- GRanges(seqnames = gene,  IRanges(start, end))
  param <- ScanBamParam(which = which, what=scanBamWhat())
  Bam <- scanBam(file = bam, index = index, param = param)
  IRanges <- IRanges(start = Bam[[1]][["pos"]], width=Bam[[1]][["qwidth"]])
  TempCov <- coverage(IRanges)
  position <- NA
  coverage <- NA
  if (length(TempCov) > 0){
    Peaks <- slice(TempCov, 0)
    position <- seq(start(Peaks[1]), end(Peaks[1]))
    coverage <- as.integer(Peaks[[1]])
  }
  id <- gsub("data/", "", bam)
  id <- gsub(".bam", "", id)
  out <- cbind.data.frame(position, coverage, id)
  colnames(out) <- c("Position", "Coverage", "Sample")
  return(out)
}
```

Load the coverage data associated with a specific gene. Change the `gene` variable to look at another gene.
```{r load_bam}
gene <- "MA_ERM_03"

AAV.3.12 <- extcov(bam = "data/AAV-3-12.bam", index = "data/AAV-3-12", gene = gene)  
```

Combine the data into a nice plot. I use the `annotate` function of `ggplot2` to mark the position of the resistance gene.
```{r plot1, fig.height=6, fig.width=10}
gs <- subset(gpos, Gene == gene)
ah <- max(AAV.3.12$Coverage)/20

ggplot(AAV.3.12, aes(x = Position, y = Coverage, color = Sample)) + 
  geom_line(lwd = 1) + 
  annotate("rect", xmin = gs$Start, xmax = gs$End, ymin = -ah-ah/10, ymax = -ah/10,  alpha = .3) +
  annotate("text", x = gs$Start + (gs$End - gs$Start)/2, y = -ah/10-ah/2, label = gene) 
```
